---
title: "p8105_hw3_kb2977"
author: "Francois Ban"
date: "10/7/2018"
output: github_document
---

##Problem 1

###Part 1: Loading brfss_data and cleaning

1. Format the data to use appropriate variable names
2. Focus on the “Overall Health” topic
3. Include only responses from “Excellent” to “Poor” (i.e. no pre-collapsed categories)
4. Organize responses as a factor taking levels from “Excellent” to “Poor”
5. Getting rid of unnecesary variables

```{r brfss_data}
library(tidyverse)

devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(brfss_smart2010)

brfss_data = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health" & response == "Excellent" | response == "Poor" | response == "Very good" | response == "Good" | response == "Fair") %>%
  mutate(response = as.factor(response)) %>% 
  select(-class, -topic, -question, -sample_size, -confidence_limit_low:-geo_location)
```

###Part 2: Answering following questions

Using the dataset above:

* In 2002, which states were observed at 7 locations?

```{r brfss_data_2002}
brfss_data_2002 = 
  brfss_data %>%
  spread(key = response, value = data_value) %>%
  filter(year == "2002") %>%
  count(locationabbr) %>% 
  filter(n == 7)
brfss_data_2002
```

* Below is a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r state_plot}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "100%"
  )

state_plot = brfss_data %>% 
  group_by(locationabbr, year) %>% 
  summarize(n_obs = n()) %>%
  ggplot(aes(x = year, y = n_obs, color = locationabbr)) +
    geom_line() +
    labs(
        title = "Observations in Each State",
        x = "Time in Years",
        y = "Number of Observations",
        caption = "Data from the brfss package"
    ) +
    scale_color_hue(name = "States")
state_plot
```

* Below is a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r table_data}
table_data = 
  brfss_data %>%
  filter(year == "2002" | year == "2006" | year == "2010") %>% 
  filter(response == "Excellent") %>%
  filter(locationabbr == "NY") %>%
  group_by(year) %>%
  summarize(mean_ex = mean(data_value),
            sd_ex = sd(data_value)) %>%
  knitr::kable(digits = 1)
```

* Below is a five-panel plot that shows, for each response category separately, the distribution of state-level averages over time.

```{r response_plot}
library(patchwork)
library(ggridges)

av_prop_plot =
  brfss_data %>% 
  spread(key = response, value = data_value) %>% 
  group_by(year, locationabbr) %>% 
  summarize(av_prop_excellent = mean(Excellent),
            av_prop_verygood = mean(`Very good`),
            av_prop_good = mean(Good),
            av_prop_fair = mean(Fair),
            av_prop_poor = mean(Poor)) %>% 
  gather(key = response, value = mean, av_prop_excellent:av_prop_poor) %>% 
  separate(response, into = c("remove_1", "remove_2", "response"), sep = "_") %>% 
  select(-remove_1,-remove_2) %>% 
  mutate(response = forcats::fct_relevel(response, c("excellent", "verygood", "good", "fair", "poor"))) %>% 
  ggplot(aes(x = year, y = mean)) +
    geom_violin(aes(), color = "blue", alpha = .5) +
    facet_grid(~response) +
    labs(
        title = "Distribution of State-Level Responses",
        x = "Time in years",
        y = "Average proportion of responses",
        caption = "Data from the rnoaa package"
    )
av_prop_plot
```

##Problem 2

```{r instacart_data}
data(instacart)
```

###Part 1: Exploration of instacart dataset

Directions: Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations. 

* The following variables were included in this dataset: `r ls(instacart)`.
* The dataset consists of `r nrow(instacart)` rows (observations) and `r ncol(instacart)` columns (variables).
* The dataset included useful information about the customer, the order, and the products of interest.  
* There were `r count(distinct(instacart, order_id))` orders placed.
* The average number of items ordered in this dataset was `r instacart %>% count(order_id) %>% summarize(ave_num_item = mean(n))`.
* The user_id with the most orders was `r names(which.max(table(instacart$user_id)))`.
* The most ordered item was `r names(which.max(table(instacart$product_name)))`.
* Customers tend to place orders around the `r names(which.max(table(instacart$order_hour_of_day)))` hour of the day.

###Part 2: Answering questions using instacart_data

* How many aisles are there, and which aisles are the most items ordered from? There were a total of `r count(distinct(instacart, aisle_id))` aisles. 

* The following aisles had the most items ordered:
`r instacart %>% count(aisle) %>% top_n(5) %>% knitr::kable(digits = 1)`

* Below is a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
mutate(aisle_id = forcats::fct_reorder(aisle_id, desc(n)))
```{r}
items_ordered = 
  instacart %>% 
  count(aisle) %>%
  mutate(aisle = forcats::fct_reorder(aisle, desc(n))) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size = 7)) +
  labs(
      title = "Plot of Items Ordered in Each Aisle",
      x = "Aisle Name",
      y = "Number of Items",
      caption = "Data from the instacart package"
  )
items_ordered
```

* Below is a table showing the top 3 popular items in the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r pop_item}
pop_item = 
  instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  top_n(3) %>% 
  knitr::kable(digits = 1)
pop_item
```

* Below is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
mean_hour_table = 
  instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
mutate(order_dow = recode_factor(order_dow, "0" = "Sunday", 
                             "1" = "Monday", 
                             "2" = "Tuesday", 
                             "3" = "Wednesday", 
                             "4" = "Thursday", 
                             "5" = "Friday", 
                             "6" = "Saturday"
                             )
         ) %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_hr = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hr) %>% 
  knitr::kable(digits = 1)
mean_hour_table
```

##Problem 3

```{r ny_noaa}
data(ny_noaa)
```

###Part 1: Exploration of ny_noaa dataset

Directions: Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. 

* The following variables were included in this dataset: `r ls(ny_noaa)`.
* The dataset consists of `r nrow(ny_noaa)` rows (observations) and `r ncol(ny_noaa)` columns (variables).
* The dataset included useful information about the weather station, date, temperature, precipitation, and snowfall.
* There were `r count(distinct(ny_noaa, date))` dates in which weather was observed.
* Weather information was gathered from `r count(distinct(ny_noaa, id))` distinct stations.
* Because weather stations collect only a subset of variables, there is a lot of missing data. We need to keep this in mind when creating graphs that describe the datset as it will get rid of any rows containing non-finite values (i.e. NA).

###Part 2: Answering questions using ny_noaa datset

* Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units.

```{r ny_noaa_clean}
ny_noaa_clean = 
  ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(prcp = as.numeric(prcp)/10,
         tmax = as.numeric(tmax)/10,
         tmin = as.numeric(tmin)/10
         )
```

For snowfall, the most commonly observed value was `r names(which.max(table(ny_noaa_clean$snow)))`. This is because most days of the year, it does not snow.

* Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r ave_temp}

ave_temp = 
  ny_noaa_clean %>% 
  filter(month == "01" | month == "07") %>% 
  mutate(month = recode(month, "01" = "January", "07" = "July")) %>% 
  group_by(month, id, year) %>% 
  summarize(ave_max_temp = mean(tmax)) %>% 
  ggplot(aes(x = year, y = ave_max_temp)) +
  geom_hex() +
  facet_grid(~month) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(size = 9)) +
  labs(
      title = "Average Maximum Temperature in January and July",
      x = "Time in Years",
      y = "Average Maximum Temperature (Celcius)",
      caption = "Data from the ny_noaa package"
  ) 
ave_temp
```

When looking at the hexigon plot created above, it is evident that the average maximum temperature for January is centered around 0 degrees Celcius across years and whereas for July, it seems to be centered around 27 degrees Celcius. In both months, there are a couple of outliers (1982 in January and 1988 in July) in which the average maximum temperature dips and is very low compared to the rest of the data.

* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r two_plots}
devtools::install_github("thomasp85/patchwork")
temp_plot = 
  ny_noaa_clean %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex() +
  labs(
      title = "Plot of Maximum vs. Minimum Temperature",
      x = "Maximum Temperature in Celcius",
      y = "Minimum Temperature in Celcius",
      caption = "Data from the ny_noaa package"
  ) 

snowfall_plot = 
  ny_noaa_clean %>% 
  filter(snow < 100 & snow > 0) %>% 
  ggplot(aes(x = snow, y = year)) +
  geom_density_ridges(alpha = .5, color = "blue", scale = 0.85) +
  labs(
      title = "Distribution of Snowfall by Year",
      x = "Snowfall (mm)",
      y = "Time in Years",
      caption = "Data from the ny_noaa package"
  ) 

temp_plot + snowfall_plot
```
